---
title: "Masters Thesis"
author: "Nick Zinck"
date: "May 2017"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
bibliography: "F:/Nick Zinck/Shiny Water Quality/Thesis/Thesis.bib"

nocite: | 
  @noauthor_shiny_nodate, @noauthor_shiny_nodate-1
---

```{r loadpack, include=FALSE}
# Load Packages
library(rmarkdown)
library(knitr)
library(tidyverse)
library(shiny)
library(kableExtra)
require(DiagrammeR)
require(DiagrammeRsvg)
require(rsvg)
```

```{r enter data,include=FALSE}
pdf_digraph <- function(filename, code, width = 500, height = NULL){
  capture.output({
  g <- grViz(paste("digraph{", code, "}"))
  DiagrammeRsvg::export_svg(g) %>% charToRaw %>% rsvg::rsvg_pdf(filename, width = width, height = height)
  },  file='NUL')
  knitr::include_graphics(filename)
}
```


# Shiny Developer Guide

## R
### Packages
## RStudio
## Shiny
### What is a Shiny App
### User Interface
### Server
### Reactivity
### Capabilities
### Limitations
## Github

Moving and renaming files should be done within Github.
If one does it in the window's explorer, Github will think this is a new file and that the previous version is deleted. Thus, for a track record, it is best to rename and move files via Github. Heere is a link for help.


https://github.com/blog/1436-moving-and-renaming-files-on-github


## Database
## Remote Connection
## Tidy Data



# Developer Guide for WAVE and WIT

## R

## RStudio

## Shiny


The DWSPA App is created with RStudio's Shiny free software. Shiny is a package in R that enables an App creator to use the R programming language to create an application that consists of a front-end HTML and JavaScript user interface in which the user interacts with the data. Shiny video and written tutorials [@____] are a great resource on how Shiny works. The Application fetches data that is in the Access Databases stored on the internal network. Upon starting the App, the App will fetch the latest data that is in the databases. The App Code is stored on Github which is an open source code sharing site. The latest Shiny Application code on Github is ran every time the App is opened. The Code that is on Github should be the most up-to-date working code that the App builders have sent to Github.

Shiny's components that the user can see consists of two categories: Input objects and Output Objects. Input objects essentially allow the user to communicate with the App and the  output objects are the App communicate with the User. input objects will be refered to as Input Widgets in this user manual and can take the form of checkboxs, selection drop-down lists, and Date Range Calendar inputs. The full list of Shiny Widgets can be seen in the Shiny Widget Gallery (@) . Output objects include plots, tables, maps, images, and tables. Shiny takes the input of from the user and executes code which are essentially directions for a sequence of computations and other operations to create the outputs that the user sees.

The Shiny Widget Gallery shows all of the available widgets in Shiny in an operational form. The 
### What is a Shiny App

### User Interface

### Server

### Reactivity

### Capabilities

### Limitations

## Github

## Database

### Tributary Data

### Reservoir Data

### Site Location Data

### Parameter Data

## Remote Connection

## Tidy Data

## R Packages

### Packages

See Section __ for Shiny Package information

### tidyverse

#### dplyr

#### tidyr

#### ggplot

#### leaflet

#### DT

### Plotly

### stringr

### scales

### RMarkdown

### knitr

### knitr

### Leaflet

### js



# References