---
title: "Masters Thesis"
author: "Nick Zinck"
date: "May 2017"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
bibliography: "F:/Nick Zinck/Shiny Water Quality/Thesis/Thesis.bib"

nocite: | 
  @noauthor_shiny_nodate, @noauthor_shiny_nodate-1
---

```{r loadpack, include=FALSE}
# Load Packages
library(rmarkdown)
library(knitr)
library(tidyverse)
library(shiny)
library(kableExtra)
require(DiagrammeR)
require(DiagrammeRsvg)
require(rsvg)
```

```{r enter data,include=FALSE}
pdf_digraph <- function(filename, code, width = 500, height = NULL){
  capture.output({
  g <- grViz(paste("digraph{", code, "}"))
  DiagrammeRsvg::export_svg(g) %>% charToRaw %>% rsvg::rsvg_pdf(filename, width = width, height = height)
  },  file='NUL')
  knitr::include_graphics(filename)
}
```


# Shiny Developer Guide

## R
### Packages
## RStudio
## Shiny
### What is a Shiny App
### User Interface
### Server
### Reactivity
### Capabilities
### Limitations
## Github

Moving and renaming files should be done within Github.
If one does it in the window's explorer, Github will think this is a new file and that the previous version is deleted. Thus, for a track record, it is best to rename and move files via Github. Heere is a link for help.


https://github.com/blog/1436-moving-and-renaming-files-on-github


## Database
## Remote Connection
## Tidy Data



# Developer Guide for WAVE and WIT

## R

## RStudio

## Shiny

### What is a Shiny App

### User Interface

### Server

### Reactivity

### Capabilities

### Limitations

## Github

## Database

### Tributary Data

### Reservoir Data

### Site Location Data

### Parameter Data

## Remote Connection

## Tidy Data

## R Packages

### Packages

See Section __ for Shiny Package information

### tidyverse

#### dplyr

#### tidyr

#### ggplot

#### leaflet

#### DT

### Plotly

### stringr

### scales

### RMarkdown

### knitr

### knitr

### Leaflet

### js



# References